---
title: "Advanced Data Manipulation in R"
subtitle: "datatime and string manipulations"
# author: "Kaixin Wang"
# date: "OSCR - Summer Session C 2019"
# output: rmarkdown::github_document
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "")
```

Import data:

```{r, message = F}
library(readr)
input = read_csv("person_info.csv")
input
```

Date-time manipulation:

In R, we recommend using the `lubridate` library to for converting data in the format of `day-month-year` to a `datetime` object.

```{r, messgae = F}
library(lubridate)
dates <- c("15/02/2013", "15 Feb 13", "It happened on 15 02 '13")
dmy(dates)

mdy(input$birthday)
```

Note: 

There are similar functions for all permutations of `d`, `m` and `y`. Explicitly, all of the following
functions exist: `dmy myd ydm mdy dym ymd`.

String manipulation:

- String normalization:

```{r}
library(stringr)
str_trim(" hello world ")
str_trim(" hello world ", side = "left")
str_trim(" hello world ", side = "right")

toupper("Hello world")
tolower("Hello World")

input$City[5]
input$City[5] = str_trim(input$City[5], side = "right")
input$City[5]
```

- Approximate string matching:

```{r}
gender <- c("M", "male ", "Female", "fem.")
grepl("m", gender)
grep("m", gender)
```
